Black swans in ecological time series\\Black swans in population dynamics\\Heavy-tailed ecological time series\\Black swans in ecological dynamics\\Black swan ecological dynamics\\ \ldots

### 2014-10-17 supp.tex
Add to SOM

Our results show...

Fig.~\ref{fig:sim-gompertz}
Fig.~\ref{fig:sim-gompertz-boxplots}
Fig.~\ref{fig:sim-prob}

>>>>

Our results show that the more-complicated Gompertz model can still recapture the true value of $\nu$ when the process noise was chosen so that appropriate tail events were present (Figs~\ref{fig:sim-gompertz} and \ref{fig:sim-gompertz-boxplots}, red and green symbols in the top rows). Likewise, the other Gompertz parameters were estimated without any systematic bias (Figs~\ref{fig:sim-gompertz} and \ref{fig:sim-gompertz-boxplots}, red and green symbols). The addition of observation error caused the model to tend to underestimate the degree of heavy-tailedness, overestimate the magnitude of process noise, somewhat overestimate $\lambda$, and overestimate density dependence (blue symbols in Figs~\ref{fig:sim-gompertz} and \ref{fig:sim-gompertz-boxplots}). The overestimation of density dependence with observation error is a known phenomenon \citep{knape2012}. Fitting a model with correctly specified observation error made marginal improvements to model bias (purple symbols in Figs~\ref{fig:sim-gompertz} and \ref{fig:sim-gompertz-boxplots}).

When converting the posterior distributions of $\nu$ into Pr($\nu < 10$) the models performed well. Without observation error, and using a probability of $0.5$ as a threshold, the model would have miscategorized only one of $40$ simulations at $\nu = 5$ across a sample size of $50$ or $100$ data points (Fig.~\ref{fig:sim-prob}, upper panels). The model would have correctly categorized all cases where the process noise was not heavy tailed (indicated as ``$\nu =$ infinity'' in Fig.~\ref{fig:sim-prob}) and all cases where $\nu = 3$. With $0.2$ standard deviations of observation error, the model still categorized XX\% of cases as heavy tailed when $\nu = 5$ and all cases where $\nu = 3$. Allowing for observation error made little improvement to the detection of heavy tails (Fig.~\ref{fig:sim-prob}, lower-left vs.\ lower-right panel). Therefore, we chose to focus on the simpler model without observation error in the main text.

%Fig.~\ref{fig:sim-gompertz}
%Fig.~\ref{fig:sim-gompertz-boxplots}
%Fig.~\ref{fig:sim-prob}
